The Mass Distribution of Stellar-Mass Black Holes

### 2010-10-23 paper.tex
Major re-ordering, added conclusion.  Shipping as v3 (but figures still need work).

 \altaffiltext{1}{School of Physics and Astronomy, University of
   Birmingham, Edgbaston, Birmingham B15 2TT, UK}

\begin{abstract}
  We perform a Bayesian analysis of the mass distribution of
  stellar-mass black holes using the observed masses of 15 low-mass
  X-ray binary systems undergoing Roche lobe overflow and five
  higher-mass, wind-fed X-ray binary systems.  Using MCMC
  calculations, we model the mass distribution both
  parametrically---as a power law, exponential, gaussian, combination
  of two gaussians, or log-normal distribution---and
  non-parametrically---as histograms with varying numbers of bins.  We
  provide confidence bounds on the shape of the mass distribution in
  the context of each model and compare the models with each other by
  calculating their Bayesian evidence as supported by the
  measurements.  The mass distribution of the low-mass systems is best
  fit by a power-law, while the distribution of the complete sample is
  best fit by the exponential model.  We examine the existence of a
  ``gap'' between the most massive neutron stars and least massive
  black holes, finding that the best model (the power law) fitted to
  the low-mass systems only gives a minimum black-hole mass of 4.3
  $\Msun$ (90\% confidence), while the best model (the exponential)
  fitted to all 20 systems gives a minumum black-hole mass of 4.5 and
  5.2 $\Msun$ (90\% confidence), respectively.  We conclude that our
  sample of black hole masses provides strong evidence of a gap
  between the maximum neutron star mass and minimum black hole mass.
\end{abstract}

@@

(X-Ray binaries are important---supernova, evolutionary processes, GW,
etc.)

\citet{Bailyn1998} examined a sample of seven transient X-ray binaries
thought to contain a black hole, concluding in a Bayesian analysis
that the mass function was strongly peaked around seven solar masses.
They found strong evidence for a ``gap'' between the least massive
black hole and the upper limit for neutron star masses of 3 $\Msun$.

A more recent study \citep{Ozel2010}, also in a Bayesian framework,
largely confirmed these findings, examining 16 low-mass X-ray binary
systems containing black holes and finding a strongly peaked
distribution at $7.8 \pm 1.2 \, \Msun$.  \citet{Ozel2010} used two
models for the mass function: a Gaussian and a decaying exponential
with a minimum ``turn-on'' mass (motivated by the analytical model of
the black-hole mass function in \citet{Fryer2001}), finding in both
cases strong evidence of a mass gap.  They argue that this mass gap
cannot be explained by observational selection effects.

In this work, we use observations of 20 low- and higher-mass galactic
X-ray binaries that contain black holes to constrain the mass function
for stellar mass black holes in the galaxy.  We consider separately
the 15 low-mass X-ray binaries containing Roche-lobe-filling
secondaries with masses smaller than the black hole primary, and a
complete sample including these systems and five higher-mass, wind-fed
systems where the secondary is heavier than the black hole primary.

We use a Bayesian Markov-Chain Monte Carlo analysis to explore various
models for the black hole mass function for both samples, including
those of \citet{Ozel2010} and \citet{Bailyn1998}.  We include both
parameteric models, such as a Gaussian, and non-parameteric models
where the mass function is represented by histograms with various
numbers of bins.  After computing posterior distributions for the
model parameters, we use model selection techniques (including a new
technique for efficient reversible-jump MCMC \citep{Farr2010}) to
compare the evidence for the various models from both samples.

@@

The theoretical model from \citet{Fryer2001}, a decaying exponential,
considered for the low-mass systems in \citet{Ozel2010}, has
relatively low evidence.  We find that the lower-mass systems are best
described by a Gaussian (the other model from \citet{Ozel2010}).  We
find that the theoretical model from \citet{Fryer2001} is, however,
the preferred model for the combined sample, while a Gaussian has much
lower evidence in this case.  A model with two separate Gaussian peaks
also has high evidence for the combined sample of systems, suggesting
\fixme{differences in the distributions of lower- and high-mass X-ray
  binary systems}.

The structure of this paper is as follows \fixme{This will change
  dramatically}.

@@

The 20 X-ray binary systems on which this study is based are listed in
Table \ref{tab:sources}.  We separate the systems into 15 low-mass
systems in which the central black hole appears to be fed by
roche-lobe overflow from the secondary, and 5 higher-mass systems in
which the black hole is fed via winds (these systems all have a
secondary that appears more massive than the black hole).  The low-
and higher-mass systems undoubtedly have different evolutionary
tracks, and therefore it is reasonable that they would have different
black-hole mass distributions.  We will first analyze the 15 low-mass
systems alone (Sections \ref{sec:models} and
\ref{sec:model-selection}), and then the complete sample of 20 systems
(Section \ref{sec:higher-mass}).

In each of these systems, spectroscopic measurements of the secondary
star provide an orbital period for the system and a half-amplitude for
the secondary's velocity curve.  These measurements can be combined
into the mass function,
\begin{equation}
  \label{eq:mass-function}
  f(M) = \frac{P K^3}{2\pi G} = \frac{M \sin^3 i}{\left( 1 + q \right)^2},
\end{equation}
where $P$ is the orbital period, $K$ is the secondary's velocity
semi-amplitude, $M$ is the black hole mass, $i$ is the inclination of
the system, and $q \equiv M_2 / M$ is the mass ratio of the system.

@@

For the purposes of this paper, we adopt the following simplified
approach to the estimation of the black hole mass from the observed
data.  When an observable is well-constrained, we assume that the true
value is normally distributed about the measured value with a standard
deviation equal to the quoted observational error.  This is the case
for the mass function in all the systems we use, and for many systems'
mass ratios and inclinations.  When a large range is quoted in the
literature for an observable, we take the true value to be distributed
uniformly (for the mass ratio) or isotropically (for the inclination)
within the quoted range.  Table \ref{tab:sources} gives the assumed
distribution for the observables in the 20 systems we use.  We do not
attempt to deal with the systematic biases in the observational
determination of $f$, $q$, and $i$ in any realistic way; we are
currently investigating more realistic treatments of the errors
(including observational biases that can shift the peak of the true
mass distribution away from the ``best-fit'' mass in the
observations).  This careful treatment will appear in future work.

@@

From these assumptions, we can generate probability distributions for
the true mass of the black hole given the observations and errors via
the Monte Carlo method: drawing samples of $f$, $q$, and $i$ from the
assumed distributions and computing the mass implied by Equation
\eqref{eq:mass-function} gives samples of $M$ from the distribution
induced by the relationship in Equation \eqref{eq:mass-function}.
Mass distributions generated in this way for the systems used in this
work are shown in Figure \ref{fig:low-masses} and Figure
\ref{fig:high-masses}.  Note that systems for which $i$ is poorly
constrained have broad ``tails'' on their mass distributions.  These
mass distributions constitute the ``observational data'' we will use
in the remainder of this paper.

@@

\section{Models}
\label{sec:models}

In this section we apply a statistical analysis to various models for
the underlying mass distribution from which the low-mass systems in
Table \ref{tab:sources} were drawn.  We will treat the combined low-
and higher-mass systems in Section \ref{sec:higher-mass}.  The end
result will be the probability distribution for the parameters of each
model implied by the data from Section \ref{sec:systems} in
combination with our prior assumptions about the probability
distribution for the parameters.  Bayes' rule relates these
quantities.  For a model with parameters $\vtheta$ in the presence of
data $d$, Bayes' rule states
\begin{equation}
  \label{eq:Bayes-rule}
  p(\vtheta | d) = \frac{p(d | \vtheta) p(\vtheta)}{p(d)}.
\end{equation}
Here, $p(\vtheta|d)$, called the posterior probability distribution
function, is the probability distribution for the parameters $\vtheta$
implied by the data $d$; $p(d|\vtheta)$, called the likelihood, is the
probability of observing data $d$ given that the model parameters are
$\vtheta$; $p(\vtheta)$, called the prior, reflects our estimate of the
probability of the various model parameters in the absence of any
data; and $p(d)$, called the evidence, is an overall normalizing
constant ensuring that
\begin{equation}
  \int d\theta\, p(\vtheta|d) = 1,
\end{equation}
whence
\begin{equation}
  \label{eq:evidence-def}
  p(d) = \int d\vtheta\, p(d|\vtheta) p(\vtheta).
\end{equation}

In our context, the data are the mass distributions given in Section
\ref{sec:systems}: $d = \{ p_i(M)| i = 1, 2, \ldots, 17 \}$.  We
assume that the measurements in Section \ref{sec:systems} are
independent, so the complete likelihood is given by a product of the
likelihoods for the individual measurements.  For a model with
parameters $\vtheta$ that predicts a mass distribution $p(M|\vtheta)$
for black holes, we have
\begin{equation}
  \label{eq:likelihood-def}
  p(d|\vtheta) = \prod_i \int dM\, p_i(M) p(M|\vtheta).
\end{equation}
That is, the likelihood of an observation is the average over the
individual mass distribution implied by the observation, $p_i(M)$, of
the probability for a black hole of that mass to exist according to
the model of the mass distribution, $p(M | \vtheta)$.  We approximate
the integrals as averages of $p(M|\vtheta)$ over the Monte Carlo mass
samples drawn from the distributions in Table \ref{tab:sources} (also
see Figures \ref{fig:low-masses} and \ref{fig:high-masses}):
\begin{equation}
  p(d|\vtheta) \approx \prod_i \frac{1}{N_i} \sum_{j = 1}^{N_i} p(M_{ij} | \vtheta),
\end{equation}
where $M_{ij}$ is the $j$th sample (out of a total $N_i$) from the
$i$th individual mass distribution.

Our calculation of the likelihood of each observation does not include
any attempt to account for selection effects in the observations.  We
simply assume (almost certainly incorrectly) that any black hole drawn
from the underlying mass distribution is equally likely to be
observed.  (See also \citet{Ozel2010}.)

For a mass distribution with several parameters, $p(\vtheta | d)$
lives in a multi-dimensional space.  Exploring the entirety of this
space systematically rapidly becomes prohibitive as the number of
parameters increases.  A more efficient way to explore the
distribution $p(\vtheta | d)$ is to use a Markov Chain Monte Carlo
(MCMC) method.  MCMC methods produce a chain (sequence) of parameter
samples, $\{ \vtheta_i, i = 1, \ldots \}$, such that a particular
parameter set, $\vtheta$, appears in the sequence with a frequency
equal to its probability according to $p(\vtheta|d)$.  In this way,
regions of parameter space where $p(\vtheta|d)$ is large are sampled
densely while regions where $p(\vtheta|d)$ is small are effectively
ignored.  A Markov chain has the property that the transition
probability from one element to the next, $p(\vtheta_i \to
\vtheta_{i+1})$, depends only on the value of $\vtheta_i$, not on any
previous values in the chain.

One way to produce a sequence of MCMC samples is via the following
algorithm, first proposed by \citet{Metropolis1953}:
\begin{enumerate}
  \item Begin with the current sample, $\vtheta_i$.
  \item Propose a new sample, $\vtheta_p$, by drawing randomly from a
    ``jump proposal distribution'' with probability $Q(\vtheta_i \to
    \vtheta_p)$.  Note that $Q(\vtheta_i \to \vtheta_p)$ can depend on
    the current parameters, $\vtheta_i$, and any other ``constant''
    data, but cannot examine the history of the chain beyond the most
    recent point.  This is necessary to preserve the Markovian
    property of the chain.
  \item Compute the ``acceptance'' probability,
    \begin{equation}
      \label{eq:paccept}
      p_{\textnormal{accept}} \equiv
      \frac{p(\vtheta_p|d)}{p(\vtheta_i|d)} \frac{Q(\vtheta_p \to
        \vtheta_i)}{Q(\vtheta_i \to \vtheta_p)}
    \end{equation}
  \item With probability $\min(1,p_{\textnormal{accept}})$ ``accept''
    the propsed $\vtheta_p$, setting $\vtheta_{i+1} = \vtheta_p$;
    otherwise set $\vtheta_{i+1} = \vtheta_i$.
\end{enumerate}
This algorithm is more likely to accept a proposed jump when it
increases the posterior (the first factor in Equation
\eqref{eq:paccept}) and when it is to a location in parameter space
from which it is easy to return (the second factor in Equation
\eqref{eq:paccept}); the combination of these influences in Equation
\eqref{eq:paccept} ensures that the equilibrium distribution of the
chain is $p(\vtheta|d)$.  As $i \to \infty$ the samples $\vtheta_i$
are distributed according to $p(\vtheta|d)$.

Once we have a chain of samples from $p(\vtheta|d)$---a distribution
of distributions---the distribution for any quantity of interest can
be computed.  For example, Figure \ref{fig:dists} shows the median
values, and the 10\% and 90\% quantiles for the value of the BH mass
probability distribution at various masses.  (Recall that we are only
using the 15 low-mass observations in this Section's analysis.)  Such
a figure can be generated from the MCMC output by computing the value
of the parameterized BH mass distribution at various masses for each
parameter sample output by the MCMC, and then plotting the quantiles
at each mass.

\begin{figure}
  \begin{center}
    \plottwo{plots/dist-parametric}{plots/dist-non-parametric}
  \end{center}
  \caption{\label{fig:dists} The median values of the black hole mass
    distribution, $p(M|\theta)$, at various masses implied by the
    posterior $p(\theta|d)$ for the models discussed in Sections
    \ref{sec:parametric-models} and
    \ref{sec:non-parametric-models}.  These distributions use only
    the 15 low-mass observations in Table \ref{tab:sources} (the
    complete sample is analyzed in Section \ref{sec:higher-mass}).
    Error bars span the 10\% to 90\% range.  Note that these
    ``distributions of distributions'' are not necessarily normalized,
    and need not be ``shaped'' like the underlying model
    distributions.}
\end{figure}

It is also common to look at the one-dimensional distribution for a
single parameter obtained by integrating over all other dimensions in
parameter space; this is approximated by a histogram of the MCMC
sample values for that parameter.  Such a distribution is called the
``marginalized'' distribution.

@@

In this subsection, we discuss the various parametric models of the
underlying black hole mass distribution that we have analyzed using
the low-mass data from Section \ref{sec:systems}.  (Analysis of these
models including the higher-mass systems appears in Section
\ref{sec:higher-mass}.)  Table \ref{tab:low-mass-parametric} gives
quantiles of the marginalized parameter distributions of the
parametric models.

\begin{table}
  \begin{center}
    \begin{tabular}{|l|c|c|c|c|c|c|}
      \hline
      Model & Parameter & 5\% & 15\% & 50\% & 85\% & 95\% \\
      \hline \hline
      Power Law (Equation \eqref{eq:power-law-dist}) & $\Mmin$ & 
      1.2786 &  4.1831 &  6.1001 &  6.5011 &  6.6250 \\
      \hline
       & $\Mmax$ & 8.5578 &  8.9214 & 23.3274 & 36.0002 & 38.8113 \\
       \hline
       & $\alpha$ & -12.4191 & -10.1894 & -6.3861 &  2.8476 &  5.6954 \\
       \hline \hline
       Decaying Exponential (Equation \eqref{eq:exp-def}) & $\Mmin$ & 
       5.0185 &  5.4439 &  6.0313 &  6.3785 &  6.5316 \\
       \hline
       & $M_0$ & 0.7796 &  0.9971 & 1.5516 &  2.4635 &  3.2518 \\
       \hline \hline
       Gaussian (Equation \eqref{eq:gaussian-def}) & $\mu$ & 
       6.6349 &  6.9130 &  7.3475 & 7.7845 & 8.0798 \\
       \hline
       & $\sigma$ & 0.7478 &  0.9050  & 1.2500 &  1.7335 & 2.1134 \\
       \hline \hline
       Two Gaussian (Equation \eqref{eq:two-gaussian-def}) & $\mu_1$ & 
       5.4506 &  6.3877 &  7.1514 &  7.6728  & 7.9803 \\
       \hline
       & $\mu_2$ & 7.2355 &  7.7387 & 12.3986 & 25.2456 & 31.4216 \\
       \hline
       & $\sigma_1$ & 0.3758 &  0.7626 &  1.2104 &  1.7981 &  2.3065 \\
       \hline
       & $\sigma_2$ & 0.2048 & 0.6421 & 1.9182 &  5.2757  & 7.2625 \\
       \hline
       & $\alpha$ & 0.0983 &  0.3526 & 0.8871 &  0.9792 &  0.9936 \\
       \hline \hline
       Log Normal (Equation \eqref{eq:log-normal-def}) & $\langle M \rangle$ & 
       6.7619 &  7.0122 &  7.4336  &  7.9159  &  8.2942 \\
       \hline 
       & $\sigma_M$ & 0.7292  &  0.8920  & 1.2704  &  1.8695  &  2.4069 \\
       \hline
    \end{tabular}
  \end{center}
  \caption{\label{tab:low-mass-parametric} Quantiles of the
    marginalized distribution for each of the parameters in the models
    discussed in Section \ref{sec:parametric-models}.  We indicate
    the 5\%, 15\%, 50\% (median), 85\%, and 95\% quantiles.  The
    marginalized distribution can be misleading when there are strong
    correlations between variables.  For example, while the
    marginalized distributions for the power law parameters are quite
    broad, the distribution of mass distributions implied by the power
    law MCMC samples is similar to the other models.  This occurs in
    spite of the broad marginalized distributions because of the
    correlations between the slope and limits of the power law
    discussed in Section \ref{sec:power-law}.}
\end{table}

@@

Many astrophysical distributions are power laws.  Let us assume that
the BH mass distribution is given by
\begin{equation}
  \label{eq:power-law-dist}
  p(M|\vtheta) = p(M|\{\Mmin, \Mmax, \alpha \}) =
  \begin{cases}
    A M^\alpha & \Mmin \leq m \leq \Mmax \\
    0 & \textnormal{otherwise}
  \end{cases}.
\end{equation}
The normalizing constant $A$ is given by 
\begin{equation}
  A = \frac{1+\alpha}{\Mmax^{1+\alpha} - \Mmin^{1+\alpha}}.
\end{equation}
We choose uniform priors on $\Mmin$ and $\Mmax \geq \Mmin$ between 0 and
$40 \Msun$, and uniform priors on the exponent $\alpha$ in a broad
range between $-15$ and $13$:
\begin{equation}
  p(\vtheta) = p(\{\Mmin, \Mmax, \alpha\}) = 
  \begin{cases}
    2 \frac{1}{40^2}\frac{1}{28} & 0 \leq \Mmin \leq \Mmax
    \leq 40, \quad -15 \leq \alpha \leq 13 \\
    0 & \textnormal{otherwise}
  \end{cases}.
\end{equation}

Our MCMC analysis output is a list of $\{\Mmin, \Mmax, \alpha\}$
values distributed according to the posterior 
\begin{equation}
  p(\vtheta|d) = p(\{\Mmin, \Mmax, \alpha\}|d) \propto p(d|\{\Mmin,
  \Mmax, \alpha\}) p(\{\Mmin, \Mmax, \alpha\}),
\end{equation}
with the likelihood $p(d|\{\Mmin, \Mmax, \alpha\})$ defined in
Equation \eqref{eq:likelihood-def}.  Figure \ref{fig:dists} presents
the resulting distribution of mass distributions inferred from the
parameters output by the MCMC.

In Figure \ref{fig:power-law} , we display a histogram of the resulting
samples in each of the parameters $\Mmin$, $\Mmax$, and $\alpha$
individually; this represents the one-dimensional ``marginalized''
distribution
\begin{equation}
  \label{eq:alpha-pdf}
  p(\alpha|d) = \int d\Mmin\, d\Mmax\, p(\{\Mmin, \Mmax, \alpha\}|d),
\end{equation}
and similarly for $\Mmin$ and $\Mmax$.

@@

\citet{Fryer2001} suggest that the black-hole mass distribution may be
well-represented by a decaying exponential with a minimum mass:
\fixme{Vicky, do you want to say anything more here?}
\begin{equation}
  \label{eq:exp-def}
  p(M|\vtheta) = p(M|\{\Mmin, M_0\}) = 
  \begin{cases}
    \frac{e^{\frac{\Mmin}{M_0}}}{M_0} \exp \left[ - \frac{M}{M_0}
    \right] & M \geq \Mmin \\
    0 & \textnormal{otherwise}
  \end{cases}.
\end{equation}
We choose a prior for this model where $\Mmin$ is uniform between 0
and $40 \Msun$.  For each $\Mmin$, we choose $M_0$ uniformly within a
range ensuring that $40\Msun$ is at least two scale masses above the
cutoff: $40\Msun \geq \Mmin + 2 M_0$.  This ensures that the majority
of the mass probability lies in the range $0 \leq M \leq 40\Msun$.
The resulting prior is
\begin{equation}
  p(\vtheta) = p(\{\Mmin, M_0\}) = 
  \begin{cases}
    \frac{4}{40^2} & 0 \leq \Mmin \leq 40, \quad 0 < M_0, \quad \Mmin
    + 2 M_0 \leq 40, \\
    0 & \textnormal{otherwise}
  \end{cases}
\end{equation}

>>>>

 \altaffiltext{1}{Also: School of Physics and Astronomy, University of
   Birmingham, Edgbaston, Birmingham B15 2TT, UK}

\begin{abstract}
  We perform a Bayesian analysis of the mass distribution of
  stellar-mass black holes using the observed masses of 15 low-mass
  X-ray binary systems undergoing Roche lobe overflow and five
  higher-mass, wind-fed X-ray binary systems.  Using Markov Chain
  Monte Carlo calculations, we model the mass distribution both
  parametrically---as a power law, exponential, gaussian, combination
  of two gaussians, or log-normal distribution---and
  non-parametrically---as histograms with varying numbers of bins.  We
  provide confidence bounds on the shape of the mass distribution in
  the context of each model and compare the models with each other by
  calculating their relative Bayesian evidence as supported by the
  measurements.  The mass distribution of the low-mass systems is best
  fit by a power-law, while the distribution of the complete sample is
  best fit by the exponential model.  We examine the existence of a
  ``gap'' between the most massive neutron stars and least massive
  black holes, finding that the best model (the power law) fitted to
  the low-mass systems only gives a minimum black-hole mass of 4.3
  $\Msun$ (90\% confidence), while the best model (the exponential)
  fitted to all 20 systems gives a minumum black-hole mass of 4.5
  $\Msun$ (90\% confidence).  We conclude that our sample of black
  hole masses provides strong evidence of a gap between the maximum
  neutron star mass and minimum black hole mass.
\end{abstract}

@@

The most massive stars probably end their lives with a supernova
explosion, becoming stellar-mass black holes.  The mass distribution
of such black holes can provide important clues to the end stages of
evolution of these stars.  In addition, the mass distribution of
stellar-mass black holes is an important input in calculations of
rates of gravitational wave emission events from coalescing neutron
star-black hole and black hole-black hole binaries in the LIGO
gravitational wave observatory \citep{Abadie2010}.

Observations of low-mass X-ray binaries in both the optical and X-ray
can provide a measurement of the mass of the compact object in these
systems.  There are now enough such measurements where the mass of the
compact object is large enough to classify it as a black hole to
attempt to statistically reconstruct the mass distribution of these
stellar-mass black holes.

The first study of the mass distribution of stellar-mass black holes,
in \citet{Bailyn1998}, examined a sample of seven transient X-ray
binaries thought to contain a black hole, concluding in a Bayesian
analysis that the mass function was strongly peaked around seven solar
masses.  \citet{Bailyn1998} found evidence of a ``gap'' between the
least massive black hole and the upper limit for neutron star masses
of 3 $\Msun$ \citep{Kalogera1996}.  Such a gap is puzzling in light of
theoretical studies that predict a continuous distribution of compact
object supernova remnant masses with a smooth transition from neutron
stars to black holes \citep{Fryer2001}.

A more recent study \citep{Ozel2010}, also in a Bayesian framework,
largely confirmed these findings, examining 16 low-mass X-ray binary
systems containing black holes and finding a strongly peaked
distribution at $7.8 \pm 1.2 \, \Msun$.  \citet{Ozel2010} use two
models for the mass function: a Gaussian and a decaying exponential
with a minimum ``turn-on'' mass (motivated by the analytical model of
the black-hole mass function in \citet{Fryer2001}).  \citet{Ozel2010}
do not provide confidence limits for the minimum black hole mass,
instead discussing only the maximum-posterior parameters for their
models.  Nevertheless, it appears that their analysis confirms the
existence of a mass gap.  \citet{Ozel2010} discuss possible selection
effects that could lead to the appearance of a mass gap, but conclude
these effects could not produce the observed gap, which they therefore
claim is a real property of the black hole mass distribution.

In this work, we use observations of 20 galactic X-ray binaries that
contain black holes to constrain the mass function for stellar mass
black holes.  We consider separately 15 low-mass X-ray binaries
containing Roche-lobe-filling secondaries with masses smaller than the
black hole primary, and a sample including these systems and five
higher-mass, wind-fed systems where the secondary is heavier than the
black hole primary.

We use a Bayesian Markov-Chain Monte Carlo analysis to explore various
models for the black hole mass function for both samples.  We include
both parameteric models, such as a Gaussian, and non-parameteric
models where the mass function is represented by histograms with
various numbers of bins.  (Our set of models includes those of
\citet{Ozel2010} and \citet{Bailyn1998}.)  After computing posterior
distributions for the model parameters, we use model selection
techniques (including a new technique for efficient reversible-jump
MCMC \citep{Farr2010}) to compare the evidence for the various models
from both samples.

@@

The theoretical model from \citet{Fryer2001}, a decaying exponential,
considered for the low-mass systems in \citet{Ozel2010}, is strongly
disfavored by our model selection.  We find that the lower-mass
systems are best described by a power law, followed closely by a
Gaussian (which is the other model from \citet{Ozel2010}).  We find
that the theoretical model from \citet{Fryer2001} is, however, the
preferred model for the combined sample of lower- and high-mass
systems.  A model with two separate Gaussian peaks also has high
evidence for the combined sample of systems.  While the failure of the
theoretically-motivated model to explain the lower-mass, Roche
lobe-filling systems suggests gaps in our understanding of the
formation of these systems, it is encouraging for population synthesis
studies \fixme{Vicky, do you have a couple of go-to papers for
  pop. synth we could cite here?} that the model of \citet{Fryer2001}
is a good fit to the full sample including both lower- and high-mass
systems.

@@

The 20 X-ray binary systems on which this study is based are listed in
Table \ref{tab:sources}.  We separate the systems into 15 low-mass
systems in which the central black hole appears to be fed by
roche-lobe overflow from the secondary, and 5 higher-mass systems in
which the black hole is fed via winds (these systems all have a
secondary that appears to be more massive than the black hole).  The
low- and higher-mass systems undoubtedly have different evolutionary
tracks, and therefore it is reasonable that they would have different
black-hole mass distributions.  We will first analyze the 15 low-mass
systems alone (Section \ref{sec:results-lower-mass}), and then the
complete sample of 20 systems (Section \ref{sec:higher-mass}).

In each of these systems, spectroscopic measurements of the secondary
star provide an orbital period for the system and a semi-amplitude for
the secondary's velocity curve.  These measurements can be combined
into the mass function,
\begin{equation}
  \label{eq:mass-function}
  f(M) = \frac{P K^3}{2\pi G} = \frac{M \sin^3 i}{\left( 1 + q \right)^2},
\end{equation}
where $P$ is the orbital period, $K$ is the secondary's velocity
semi-amplitude, $M$ is the black hole mass, $i$ is the inclination of
the system, and $q \equiv M_2 / M$ is the mass ratio of the system.

@@

For the purposes of this paper, we adopt the following simplified
approach to the estimation of the black hole mass from the observed
data.  When an observable is well-constrained, we assume that the true
value is normally distributed about the measured value with a standard
deviation equal to the quoted observational error.  This is the case
for the mass function in all the systems we use, and for many systems'
mass ratios and inclinations.  When a large range is quoted in the
literature for an observable, we take the true value to be distributed
uniformly (for the mass ratio) or isotropically (for the inclination)
within the quoted range.  Table \ref{tab:sources} gives the assumed
distribution for the observables in the 20 systems we use.  We do not
attempt to deal with the systematic biases in the observational
determination of $f$, $q$, and $i$ in any realistic way; we are
currently investigating more realistic treatments of the errors
(including observational biases that can shift the peak of the true
mass distribution away from the ``best-fit'' mass in the
observations).  This treatment will appear in future work.

@@

From these assumptions, we can generate probability distributions for
the true mass of the black hole given the observations and errors via
the Monte Carlo method: drawing samples of $f$, $q$, and $i$ from the
assumed distributions and computing the mass implied by Equation
\eqref{eq:mass-function} gives samples of $M$ from the distribution
induced by the relationship in Equation \eqref{eq:mass-function}.
Mass distributions generated in this way for the systems used in this
work are shown in Figure \ref{fig:low-masses} and Figure
\ref{fig:high-masses}.  Systems for which $i$ is poorly constrained
have broad ``tails'' on their mass distributions.  These mass
distributions constitute the ``observational data'' we will use in the
remainder of this paper.

@@

\section{Statistical Analysis}
\label{sec:models}

In this section we describe the statistical analysis we will apply to
various models for the underlying mass distribution from which the
low-mass systems in Table \ref{tab:sources} were drawn.  Results of
the analysis appear in Section \ref{sec:results}.

\subsection{Bayesian Inference}

The end result of our statistical analysis will be the probability
distribution for the parameters of each model implied by the data from
Section \ref{sec:systems} in combination with our prior assumptions
about the probability distribution for the parameters.  Bayes' rule
relates these quantities.  For a model with parameters $\vtheta$ in
the presence of data $d$, Bayes' rule states
\begin{equation}
  \label{eq:Bayes-rule}
  p(\vtheta | d) = \frac{p(d | \vtheta) p(\vtheta)}{p(d)}.
\end{equation}
Here, $p(\vtheta|d)$, called the posterior probability distribution
function, is the probability distribution for the parameters $\vtheta$
implied by the data $d$; $p(d|\vtheta)$, called the likelihood, is the
probability of observing data $d$ given that the model parameters are
$\vtheta$; $p(\vtheta)$, called the prior, reflects our estimate of the
probability of the various model parameters in the absence of any
data; and $p(d)$, called the evidence, is an overall normalizing
constant ensuring that
\begin{equation}
  \int d\theta\, p(\vtheta|d) = 1,
\end{equation}
whence
\begin{equation}
  \label{eq:evidence-def}
  p(d) = \int d\vtheta\, p(d|\vtheta) p(\vtheta).
\end{equation}

In our context, the data are the mass distributions given in Section
\ref{sec:systems}: $d = \{ p_i(M)| i = 1, 2, \ldots, 20 \}$.  We
assume that the measurements in Section \ref{sec:systems} are
independent, so the complete likelihood is given by a product of the
likelihoods for the individual measurements.  For a model with
parameters $\vtheta$ that predicts a mass distribution $p(M|\vtheta)$
for black holes, we have
\begin{equation}
  \label{eq:likelihood-def}
  p(d|\vtheta) = \prod_i \int dM\, p_i(M) p(M|\vtheta).
\end{equation}
That is, the likelihood of an observation is the average over the
individual mass distribution implied by the observation, $p_i(M)$, of
the probability for a black hole of that mass to exist according to
the model of the mass distribution, $p(M | \vtheta)$.  We approximate
the integrals as averages of $p(M|\vtheta)$ over the Monte Carlo mass
samples drawn from the distributions in Table \ref{tab:sources} (also
see Figures \ref{fig:low-masses} and \ref{fig:high-masses}):
\begin{equation}
  p(d|\vtheta) \approx \prod_i \frac{1}{N_i} \sum_{j = 1}^{N_i} p(M_{ij} | \vtheta),
\end{equation}
where $M_{ij}$ is the $j$th sample (out of a total $N_i$) from the
$i$th individual mass distribution.

Our calculation of the likelihood of each observation does not include
any attempt to account for selection effects in the observations.  We
simply assume (almost certainly incorrectly) that any black hole drawn
from the underlying mass distribution is equally likely to be
observed.  The results of \citet{Ozel2010} suggest that selection
effects are unlikely to significantly bias our analysis.

For a mass distribution with several parameters, $p(\vtheta | d)$
lives in a multi-dimensional space.  Previous works
\citep{Ozel2010,Bailyn1998} have considered models with only two
parameters, evaluating $p(\vtheta|d)$ on a grid.  Many of our models
for the underlying mass distribution have three or more parameters.
Exploring the entirety of these parameter spaces with a grid rapidly
becomes prohibitive as the number of parameters increases.  A more
efficient way to explore the distribution $p(\vtheta | d)$ is to use a
Markov Chain Monte Carlo (MCMC) method.  MCMC methods produce a chain
(sequence) of parameter samples, $\{ \vtheta_i \, | \, i = 1, \ldots
\}$, such that a particular parameter sample, $\vtheta$, appears in
the sequence with a frequency equal to its posterior probability,
$p(\vtheta|d)$.  In this way, regions of parameter space where
$p(\vtheta|d)$ is large are sampled densely while regions where
$p(\vtheta|d)$ is small are effectively ignored.

Once we have a chain of samples from $p(\vtheta|d)$, the distribution
for any quantity of interest can be computed by evaluating it on each
sample in the chain and forming a histogram of these values.  For
example, to compute the one-dimensional distribution for a single
parameter obtained by integrating over all other dimensions in
parameter space, called the ``marginalized'' distribution, one plots
the histogram of the values of that parameter appearing in the chain.

@@

Here we discuss the various parametric models of the underlying black
hole mass distribution considered in this paper.

@@

Many astrophysical distributions are power laws.  Let us assume that
the BH mass distribution is given by
\begin{equation}
  \label{eq:power-law-dist}
  p(M|\vtheta) = p(M|\{\Mmin, \Mmax, \alpha \}) =
  \begin{cases}
    A M^\alpha & \Mmin \leq m \leq \Mmax \\
    0 & \textnormal{otherwise}
  \end{cases}.
\end{equation}
The normalizing constant $A$ is 
\begin{equation}
  A = \frac{1+\alpha}{\Mmax^{1+\alpha} - \Mmin^{1+\alpha}}.
\end{equation}
We choose uniform priors on $\Mmin$ and $\Mmax \geq \Mmin$ between 0 and
$40 \Msun$, and uniform priors on the exponent $\alpha$ in a broad
range between $-15$ and $13$:
\begin{equation}
  p(\vtheta) = p(\{\Mmin, \Mmax, \alpha\}) = 
  \begin{cases}
    2 \frac{1}{40^2}\frac{1}{28} & 0 \leq \Mmin \leq \Mmax
    \leq 40, \quad -15 \leq \alpha \leq 13 \\
    0 & \textnormal{otherwise}
  \end{cases}.
\end{equation}

Our MCMC analysis output is a list of $\{\Mmin, \Mmax, \alpha\}$
values distributed according to the posterior 
\begin{equation}
  p(\vtheta|d) = p(\{\Mmin, \Mmax, \alpha\}|d) \propto p(d|\{\Mmin,
  \Mmax, \alpha\}) p(\{\Mmin, \Mmax, \alpha\}),
\end{equation}
with the likelihood $p(d|\{\Mmin, \Mmax, \alpha\})$ defined in
Equation \eqref{eq:likelihood-def}.

@@

\citet{Fryer2001} studied the relation between progenitor and remnant
mass in simulations of supernova explosions.  Combining this with the
mass function for supernova progenitors, they suggested that the
black-hole mass distribution may be well-represented by a decaying
exponential with a minimum mass: \fixme{Vicky, do you want to say
  anything more here?}
\begin{equation}
  \label{eq:exp-def}
  p(M|\vtheta) = p(M|\{\Mmin, M_0\}) = 
  \begin{cases}
    \frac{e^{\frac{\Mmin}{M_0}}}{M_0} \exp \left[ - \frac{M}{M_0}
    \right] & M \geq \Mmin \\
    0 & \textnormal{otherwise}
  \end{cases}.
\end{equation}
We choose a prior for this model where $\Mmin$ is uniform between 0
and $40 \Msun$.  For each $\Mmin$, we choose $M_0$ uniformly within a
range ensuring that $40\Msun$ is at least two scale masses above the
cutoff: $40\Msun \geq \Mmin + 2 M_0$.  This ensures that the majority
of the mass probability lies in the range $0 \leq M \leq 40\Msun$.
The resulting prior is
\begin{equation}
  p(\vtheta) = p(\{\Mmin, M_0\}) = 
  \begin{cases}
    \frac{4}{40^2} & 0 \leq \Mmin \leq 40, \quad 0 < M_0, \quad \Mmin
    + 2 M_0 \leq 40, \\
    0 & \textnormal{otherwise}
  \end{cases}
\end{equation}
